# Reproducible Research: Peer Assessment 1

```{r}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 1, digits = 2)
```


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
```



## What is mean total number of steps taken per day?
The distribution of the total number of steps per day can be seen in the following histogram.

```{r}
sum.per.day <- sapply(split(activity$steps,f=activity$date),sum)
hist(sum.per.day)
mean.sum.per.day <-  mean(sum.per.day,na.rm=T)
median.sum.per.day <-  median(sum.per.day,na.rm=T)

```


The __mean__ total number of steps per day is `r mean.sum.per.day` and the __median__ total number of steps per day is `r median.sum.per.day`



## What is the average daily activity pattern?
```{r}
mean.per.interval <- sapply(split(activity$steps,activity$interval),mean,na.rm=T)
plot(mean.per.interval,type="l",xaxt="n",xlab="interval",ylab="average steps")
labels <- seq(from=1,to=288,length.out=5)
axis(1,labels=names(mean.per.interval)[labels],at=labels)

```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
